---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(janitor)
library(tidyverse)
library(dplyr)
library(lubridate)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
population_by_age <- read_csv('sb152xr1685/population-by-age.csv') |> clean_names()
private_enrollment_by_grade <- read_csv('sb152xr1685/private-enrollment-by-grade.csv') |> clean_names()
public_enrollment_by_grade <- read_csv('sb152xr1685/public-enrollment-by-grade.csv') |> clean_names()
summary_file <- read_csv('sb152xr1685/Enrollment Changes-Table 1.csv')|>clean_names()
# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: Overall, public school enrollment fell by over 700,000 students between the 2019-2020 and 2021-2022 school years in the 21 states plus Washington, D.C., that provided the necessary data.
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
 pub_enroll_with_state_count <- public_enrollment_by_grade|>
   mutate(used_state = case_when(
    str_detect(st, "CA") ~ "true",
    str_detect(st, "NY") ~ "true",
    str_detect(st, "LA") ~ "true",
    str_detect(st, "NC") ~ "true",
    str_detect(st, "WA") ~ "true",
    str_detect(st, "CO") ~ "true",
    str_detect(st, "GA") ~ "true",
    str_detect(st, "MN") ~ "true",
    str_detect(st, "WI") ~ "true",
    str_detect(st, "NH") ~ "true",
    str_detect(st, "MA") ~ "true",
    str_detect(st, "MT") ~ "true",
    str_detect(st, "DE") ~ "true",
    str_detect(st, "ND") ~ "true",
    str_detect(st, "DC") ~ "true",
    str_detect(st, "SD") ~ "true",
    str_detect(st, "SC") ~ "true",
    str_detect(st, "NE") ~ "true",
    str_detect(st, "RI") ~ "true",
    str_detect(st, "TN") ~ "true",
    str_detect(st, "PA") ~ "true",
    str_detect(st, "FL") ~ "true",
    .default = "false"
   ))

  pub_enroll_with_state_count|>
    filter(used_state=="true")|>
    filter(grade=="k_12_total")|>
    group_by(sch_year)|>
    summarise(
      total=sum(enrollment, na.rm = TRUE))|>
    arrange(desc(total))
  
  pre_covid_publics <- pub_enroll_with_state_count |> 
  filter(sch_year == "2019-2020") |>
  filter(used_state == "true") |>
  summarize(total = sum(enrollment, na.rm = TRUE))

post_covid_publics <- pub_enroll_with_state_count |> 
  filter(sch_year == "2021-2022") |>
  filter(used_state == "true") |>
  summarize(total = sum(enrollment, na.rm = TRUE))

  
pre_covid_publics - post_covid_publics
# Display results of code below this codeblock

```

### Sentence 2

-   **Sentence text**: Those states saw private-school enrollment grow by over 100,000 students.
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
 priv_enroll_with_state_count <- private_enrollment_by_grade|>
   mutate(used_state = case_when(
    str_detect(state, "CA") ~ "true",
    str_detect(state, "NY") ~ "true",
    str_detect(state, "LA") ~ "true",
    str_detect(state, "NC") ~ "true",
    str_detect(state, "WA") ~ "true",
    str_detect(state, "CO") ~ "true",
    str_detect(state, "GA") ~ "true",
    str_detect(state, "MN") ~ "true",
    str_detect(state, "WI") ~ "true",
    str_detect(state, "NH") ~ "true",
    str_detect(state, "MA") ~ "true",
    str_detect(state, "MT") ~ "true",
    str_detect(state, "DE") ~ "true",
    str_detect(state, "ND") ~ "true",
    str_detect(state, "DC") ~ "true",
    str_detect(state, "SD") ~ "true",
    str_detect(state, "SC") ~ "true",
    str_detect(state, "NE") ~ "true",
    str_detect(state, "RI") ~ "true",
    str_detect(state, "TN") ~ "true",
    str_detect(state, "PA") ~ "true",
    str_detect(state, "FL") ~ "true",
    .default = "false"
   ))
 
   priv_enroll_with_state_count|>
    filter(used_state=="true")|>
    filter(grade=="k_12_total")|>
    group_by(sch_year)|>
    summarise(
      total=sum(enrollment, na.rm = TRUE))|>
    arrange(desc(total))
   
   pre_covid_privates <- priv_enroll_with_state_count |> 
  filter(sch_year == "2019-2020") |>
  filter(used_state == "true") |>
  summarize(total = sum(enrollment, na.rm = TRUE))

post_covid_privates <- priv_enroll_with_state_count |> 
  filter(sch_year == "2021-2022") |>
  filter(used_state == "true") |>
  summarize(total = sum(enrollment, na.rm = TRUE))

  pre_covid_privates - post_covid_privates
   
# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
homeschool <- summary_file |>
  filter(enrollment_type == "Homeschool enrollment")

homeschool_with_states <- homeschool |>
mutate(used_state = case_when(
    str_detect(state_abbrev, "CA") ~ "true",
    str_detect(state_abbrev, "NY") ~ "true",
    str_detect(state_abbrev, "LA") ~ "true",
    str_detect(state_abbrev, "NC") ~ "true",
    str_detect(state_abbrev, "WA") ~ "true",
    str_detect(state_abbrev, "CO") ~ "true",
    str_detect(state_abbrev, "GA") ~ "true",
    str_detect(state_abbrev, "MN") ~ "true",
    str_detect(state_abbrev, "WI") ~ "true",
    str_detect(state_abbrev, "NH") ~ "true",
    str_detect(state_abbrev, "MA") ~ "true",
    str_detect(state_abbrev, "MT") ~ "true",
    str_detect(state_abbrev, "DE") ~ "true",
    str_detect(state_abbrev, "ND") ~ "true",
    str_detect(state_abbrev, "DC") ~ "true",
    str_detect(state_abbrev, "SD") ~ "true",
    str_detect(state_abbrev, "SC") ~ "true",
    str_detect(state_abbrev, "NE") ~ "true",
    str_detect(state_abbrev, "RI") ~ "true",
    str_detect(state_abbrev, "TN") ~ "true",
    str_detect(state_abbrev, "PA") ~ "true",
    str_detect(state_abbrev, "FL") ~ "true",
    .default = "false"
   ))

  homeschool_with_states |>
    filter(used_state == "true")|>
    summarise(
      total = sum(total_change, na.rm = TRUE)
    )

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
